The objective of this thesis was to explore and improve the robustness of results in neuroimaging. This process
began with the development of a scalable library for the deployment, management, visualization, and
provenance-rich capture of computational experiments. This tool significantly lowered the barrier to adopting high
performance computing infrastructures for large neuroimaging experiments. With the computational backbone in place,
I compared various perturbation methods for the evaluation of numerical instabilities in structural connectome
estimation pipelines. Through perturbation of either the inputs directly or adding stochasticity to existing
numerical error, I was able to demonstrate instability in structural connectome estimation that peaked at the level
of observing cross-subject differences. Building upon this proof-of-concept for the effective perturbation of
pipelines, I performed a series of experiments, comparisons, and evaluations of perturbed structural connectomes to
quantify the analytical impact of the induced instabilities. While individual estimates of connectivity and
subsequently derived features varied considerably, in some cases retaining no signal, it became evident that the
perturbation of connectome generation led to variability that both improved the reliability of the dataset and had
potential for improving the downstream modelling of brain-phenotype relationships. In an effort to capture this
variance, I devised and tested several strategies for the aggregation of this variance across perturbed connectomes
and evaluated their impact on the performance of several classifiers tasked with modelling a relationship between
structural connectivity and age. I found that the aggregation of induced variability lead to more performant
predictive models which were more generalizable to unseen data.

I believe that my thesis clearly demonstrates the impact that numerical instabilities may have on neuroimaging
results, and provided a compelling solution to this problem that can be immediately adopted by the community.
Alongside the characterization of instability in structural connectome estimation, the tools I developed increase the
accessibility of the large scale application of numerical analysis and neuroimaging workflows. I began this thesis
with the ambition of shedding light on important issues in neuroscience, and ultimately improving the trustworthiness
of neuroimaging claims. I believe that the contributions presented here provide compelling evidence of both the
impact of numerical error in neuroimaging and provide a glimpse of a hopeful future in the field when these
challenges are faced head-on rather than swept under the rug.

